---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r set_options, echo=F, message=F}
require(knitr)
opts_chunk$set( fig.path = 'figure/' )
```

## Loading and preprocessing the data

The data is loaded and data and time objects are created.

```{r load_data}
data<-read.csv(unz("activity.zip","activity.csv"))
data$time<-as.POSIXct(strptime(paste(data$date,formatC(data$interval,digits=3,flag="0")),"%Y-%m-%d %H%M",tz="UTC"))
data$date<-as.Date(data$date,"%Y-%m-%d")
```


## What is mean total number of steps taken per day?

```{r steps_per_day}
stepPerDay<-tapply(data$steps,data$date,sum,na.rm=TRUE);
hist(stepPerDay,xlab = "Steps per day",col="red",breaks = 10,main = "Histogram of Steps per day")
abline(v=mean(stepPerDay),col="blue")
mean(stepPerDay)
median(stepPerDay)
```

The mean steps per day is `r mean(stepPerDay)` and the median is `r median(stepPerDay)`.

## What is the average daily activity pattern?

```{r daily_activiy_pattern}
library(ggplot2)
stepPerInterval<-aggregate(steps~interval,data=data,FUN=mean,na.rm=TRUE);
qplot(interval,steps,data = stepPerInterval,geom="line",ylab = "Number of Steps",xlab = "Interval",main = "Average Steps per interval")
stepPerInterval$interval[which.max(stepPerInterval$steps)]
```

The interval `r stepPerInterval$interval[which.max(stepPerInterval$steps)]` has an average of `r max(stepPerInterval$steps)`, that is the maximum interval average.

## Imputing missing values

There are only missin value on steps and they are `r sum(is.na(data$steps))`.
```{r missing_values_estimate}
summary(data)
sum(is.na(data$steps))
auxData<-aggregate(data$steps,by=list(data$interval),FUN = median,na.rm=TRUE)
names(auxData)<-c("interval","median.steps")
filledData<-merge(data,auxData)
filledData$steps[is.na(filledData$steps)]<-filledData$median.steps[is.na(filledData$steps)]
summary(filledData)

stepPerDayAdj<-tapply(filledData$steps,filledData$date,sum,na.rm=TRUE);
hist(stepPerDayAdj,xlab = "Steps per day",col="red",breaks = 10,main = "Histogram of Steps per day")
abline(v=mean(stepPerDayAdj),col="blue")
mean(stepPerDayAdj)
median(stepPerDayAdj)
```

The adjusted mean steps per day is `r mean(stepPerDayAdj)` and the median is `r median(stepPerDayAdj)`.


## Are there differences in activity patterns between weekdays and weekends?

```{r pattern_weekdays}
Sys.setlocale("LC_TIME", "C");
filledData$weekday=as.factor(ifelse(weekdays(filledData$time,abbreviate = TRUE) %in% c("Sat","Sun"),"weekday","weekend"));
filledData$weekday=relevel(filledData$weekday,"weekday")
filledDataByInterval<-aggregate(steps~interval+weekday,data=filledData,FUN = mean,na.rm=TRUE)
library(lattice)
qplot(interval,steps,data=filledDataByInterval,facets = weekday~.,geom="line",ylab = "Number of Steps",xlab = "Interval",main = "Average number os steps per Interval: Day Pattern")
```

